SunPy - Python for Solar Physics

### 2014-01-23 1-Introduction.tex
bibliography formating and updates

Currently, most solar data analysis is performed with a large library of 
routines called SolarSoft, also referred to as SSW \cite{ssw_ref}. SolarSoft is a set of integrated software 
libraries, databases, and system utilities which provide a common programming 
and data analysis environment for solar physics. It is overwhelmingly an 
environment which relies upon IDL (Interactive Data Language), a commercial and 
closed-source interactive data analysis environment sold by Exelis, a global 
aerospace, defense, information and services company. The SSW gen package
provides the base analysis capabilities such as time series analysis, time 
conversions, spectral fitting, image display, and file I/O. It is composed of 
89\% IDL code while the remainder is mostly perl (5.4\%) and csh scripts 
(3.8\%). According to data generated using David A. Wheeler's 'SLOCCount', 
SSW/gen is composed of 289,724 lines of code with a total approximate value of 
\$10.4M (based on the Basic COCOMO model). Individual missions and instruments 
provide additional optional packages to support and process their data. 
For example, the SDO package \cite{sdo} is 100\% IDL code and is 11k lines of code (\$360k) 
and the SOHO package \cite{soho} is 100k lines of 
code (\$3M) and is 99.9\% IDL code. SSW is hosted and distributed by the 
Lockheed Martin Solar and Astrophysics Laboratory.

@@

One of SunPy's key aims is to provide a free and modern alternative 
to the SolarSoft (gen) library. This is made possible by the recent rise of the 
scientific Python community. Python is a general purpose high-level programming 
language. It is free to use, cross-platform and is both an accessible and powerful
programming language. Python is one of the top ten most popular programming languages 
in the world according to the 2014 TIOBE Index\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}.
It therefore has a very large user community which extends far outside of the solar 
physics or astronomy communities. It is very extensible with C, C++, Fortran or 
even IDL. Many books and on-line documentation is 
available. For example at the time writing a search on Amazon for books 
on IDL programming generated $189$ results while a search for Python yielded 
$1,788$ results. Python has a strong emphasis on 
readability a concept which is well-aligned with the scientific endeavor.
Additionally, Python is increasingly being taught in universities,
which means that research projects can begin quickly 
without the need for students to learn a new (and relatively niche) 
language. Finally, knowledge of Python is a skill which is useful and well 
compensated outside of the scientific community (average salary of \$110k in 
the Washington DC area according to indeed.com at the time of writing).
Python continues to see increased use in the astronomy community 
\cite{2011ASPC..442..425G} which has similar goals and requirements as the solar physics
community. Finally, Python plays well with many technologies such as web servers \cite{2008JPhCS.119d2011D}, 
SQL databases, interactive and shareable notebook-based computing \cite{PER-GRA:2007}.
For all of these reasons, Python is an ideal platform for a modern solar physics 
data analysis environment.

SunPy is built on top of the core scientific Python packages namely 
\texttt{NumPy}, \texttt{SciPy}, \texttt{matplotlib}. These core packages provide
capabilities surpassing that provided by IDL for array manipulation and 2-d plotting
\cite{2011ASPC..442..425G}. Providing similar capabilities as SSW/gen in Python is made
easier by leveraging the diverse Python community. For example, SunPy makes use of 
the \texttt{Pandas} package to provide high-performance, easy-to-use data structures and data analysis tools
for time series data\cite{mckinney2012python}. \texttt{Pandas} was originally developed for quantitative analysis of financial
data and has seen grown into a generalized data analysis package. SunPy also relies on 
and is coordinating with the \texttt{Astropy} package \cite{2013A&A...558A..33A}. It is still in
active development but already provides many relevant capabilities such as full-featured
support for FITS files, absolute dates and time conversions, units and unit conversions,
and model-fitting. Pandas (134k lines of code) and AstroPy (90k lines of code) allow
SunPy to leverage a total value of \$7.6M worth of development which is close to the 
\$10.4M for SSW/gen assuming a rough equivalency between a line of Python code and a line
of IDL code.

@@

%citations
% 2013A&A...558A..33A http://adsabs.harvard.edu/abs/2013A&A...558A..33A astropy
% 2011ASPC..442..425G http://adsabs.harvard.edu/abs/2011ASPC..442..425G greenfield2011
% 2013ApJ...763...86L  http://adsabs.harvard.edu/abs/2013ApJ...763...86L chianti
%@book{mckinney2012python,
%  title={Python for Data Analysis: Data Wrangling with Pandas, NumPy, and IPython},
%  author={McKinney, W.},
%  isbn={9781449323615},
%  url={http://books.google.com/books?id=v3n4\_AK8vu0C},
%  year={2012},
%  publisher={O'Reilly Media}
%}
% 2008JPhCS.119d2011D http://adsabs.harvard.edu/abs/2008JPhCS.119d2011D rapid web development in python
%@Article{PER-GRA:2007,
%  Author    = {P\'erez, Fernando and Granger, Brian E.},
%  Title     = {{IP}ython: a System for Interactive Scientific Computing},
%  Journal   = {Computing in Science and Engineering},
%  Volume    = {9},
%  Number    = {3},
%  Pages     = {21--29},
%  month     = may,
%  year      = 2007,
%  url       = "http://ipython.org",
%  ISSN      = "1521-9615",
%  doi       = {10.1109/MCSE.2007.53},
%  publisher = {IEEE Computer Society},
%}

>>>>

Currently, most solar data analysis is performed with a large library 
of routines called SolarSoft, also referred to as SSW 
\cite{freeland1998}. SolarSoft is a set of integrated software 
libraries, databases, and system utilities which provide a common 
programming and data analysis environment for solar physics. It is 
overwhelmingly an environment which relies upon IDL (Interactive Data 
Language), a commercial and closed-source interactive data analysis 
environment sold by Exelis, a global aerospace, defense, information 
and services company. The SSW gen package provides the base analysis 
capabilities such as time series analysis, time conversions, spectral 
fitting, image display, and file I/O. It is composed of 89\% IDL code 
while the remainder is mostly perl (5.4\%) and csh scripts (3.8\%). 
According to data generated using David A. Wheeler's 'SLOCCount', 
SSW/gen is composed of 289,724 lines of code with a total approximate 
value of \$10.4M (based on the Basic COCOMO model). Individual 
missions and instruments provide additional optional packages to 
support and process their data. For example, the SDO package 
\cite{sdo} is 100\% IDL code and is 11k lines of code (\$360k) 
and the SOHO package \cite{soho} is 100k lines of code (\$3M) and is 
99.9\% IDL code. SSW is hosted and distributed by the Lockheed Martin 
Solar and Astrophysics Laboratory.

@@

One of SunPy's key aims is to provide a free and modern alternative 
to the SolarSoft (gen) library. This is made possible by the recent 
rise of the scientific Python community. Python is a general purpose 
high-level programming language. It is free to use, cross-platform 
and is both an accessible and powerful programming language. Python 
is one of the top ten most popular programming languages in the world 
according to the 2014 TIOBE Index 
\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}.
It therefore has a very large user community which extends faroutside 
of the solar physics or astronomy communities. It is very extensible 
with C, C++, Fortran or even IDL. Many books and on-line 
documentation is available. For example at the time writing a search 
on Amazon for books on IDL programming generated $189$ results while 
a search for Python yielded $1,788$ results. Python has a strong 
emphasis on readability a concept which is well-aligned with the 
scientific endeavor.
Additionally, Python is increasingly being taught in universities,
which means that research projects can begin quickly 
without the need for students to learn a new (and relatively niche) 
language. Finally, knowledge of Python is a skill which is useful and well 
compensated outside of the scientific community (average salary of 
\$110k in the Washington DC area according to indeed.com at the time 
of writing). Python continues to see increased use in the astronomy 
community \cite{greenfield2011} which has similar goals and 
requirements as the solar physics community. Finally, Python plays 
well with many technologies such as web servers \cite{dolgert2008}, 
SQL databases, interactive and shareable notebook-based computing 
\cite{perez2007}. For all of these reasons, Python is an ideal 
platform for a modern solar physics data analysis environment.

SunPy is built on top of the core scientific Python packages namely 
\texttt{NumPy}, \texttt{SciPy}, \texttt{matplotlib}. These core 
packages provide capabilities surpassing that provided by IDL for 
array manipulation and 2-d plotting \cite{greenfield2011}. Providing 
similar capabilities as SSW/gen in Python is made easier by 
leveraging the diverse Python community. For example, SunPy makes use 
of the \texttt{Pandas} package to provide high-performance, 
easy-to-use data structures and data analysis tools for time series 
data\cite{mckinney2012}. \texttt{Pandas} was originally developed for 
quantitative analysis of financial data and has seen grown into a 
generalized data analysis package. SunPy also relies on and is 
coordinating with the \texttt{Astropy} package 
\cite{theastropycollaboration2013}. It is still in active development 
but already provides many relevant capabilities such as full-featured
support for FITS files, absolute dates and time conversions, units 
and unit conversions, and model-fitting. Pandas (134k lines of code) 
and AstroPy (90k lines of code) allow SunPy to leverage a total value 
of \$7.6M worth of development which is close to the \$10.4M for 
SSW/gen assuming a rough equivalency between a line of Python code 
and a line of IDL code.

@@

### 2014-01-23 5-Dev.tex
bibliography formating and updates

\section{Development and Community}
SunPy is a community developed library, it is designed and developed for and by 
the solar physics community. Not only is all the source code publicly available 
online under the permissive 2-clause BSD licence 
\footnote{\url{http://opensource.org/licenses/BSD-2-Clause}}, the whole 
development 
process is also online and open for anyone to contribute to.
The primary media by which SunPy's development is facilitated is the online 
service GitHub \footnote{\url{http://github.com}} which makes use of the 
distributed 
version control software git. \footnote{\url{http://git-scm.com/}}

>>>>

\section{Development and Community}
SunPy is a community developed library, it is designed and developed for and by 
the solar physics community. Not only is all the source code publicly available 
online under the permissive 2-clause BSD licence 
\footnote{\url{http://opensource.org/licenses/BSD-2-Clause}}, the whole 
development 
process is also online and open for anyone to contribute to.
The primary media by which SunPy's development is facilitated is the online 
service GitHub \footnote{\url{http://github.com}} which makes use of the 
distributed 
version control software \href{http://git-scm.com/}{git}.
